---
pdf_document: default
title: "Multiple Linear Regression 1"
abstract: |
  This is the code for the first linear regression lesson.
output: pdf_document
---

```{r "setup", include=FALSE}
#This block of code will not appear in your knitted document
#Setup paths for data and Rcode
require("knitr")
traindir <- "~/Google Drive/UVA/Courses/LSM/Fall2020/TrainData/"
sourcedir <-"~/Google Drive/UVA/Courses/LSM/Fall2020/Source/"
opts_knit$set(root.dir = sourcedir)
opts_chunk$set(warning=FALSE)
```

Load data and combine data from 2001-2019 into totacts
```{r}
# Source AccidentInput
source("AccidentInput.R")

# you should have two data structures in working memory
# First - a list of data frames for each year of accident data
acts <- file.inputl(traindir)

# Next a data frame with all accidents from all years from 2001 - 2018
# with columns that are consistent for all of these years

# Get a common set the variables

comvar <- intersect(colnames(acts[[1]]), colnames(acts[[8]]))

# the combined data frame
totacts <- combine.data(acts, comvar)
```

Build a data frame with only extreme accidents for ACCDMG
```{r}
dmgbox <- dmgbox <-boxplot(totacts$ACCDMG)
  
ggplot(as.data.frame(totacts$ACCDMG), aes(x=totacts$ACCDMG)) + 
geom_boxplot(col= "steelblue") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

xdmg <- totacts[totacts$ACCDMG > dmgbox$stats[5],]
```

Data cleaning 
```{r}
# Remove 9/11
xdmg <- xdmg[-190,]

##Remove duplicates
xdmgnd <- xdmg[!(duplicated(xdmg[, c("INCDTNO", "YEAR", "MONTH", 
                                     "DAY", "TIMEHR", "TIMEMIN")])),]

# Reset rownames (observation #s) for sequential numbering- 
# otherwise they will remain the #s from totacts
rownames(xdmgnd) <- NULL
```

Setup categorical variables

This new variable, Cause, has to be a factor
```{r}
xdmgnd$Cause <- rep(NA, nrow(xdmgnd))

xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "M")] <- "M"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "T")] <- "T"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "S")] <- "S"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "H")] <- "H"
xdmgnd$Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "E")] <- "E"

xdmgnd$Cause <- factor(xdmgnd$Cause)
```

Setup the Type variable
```{r}
xdmgnd$Type <- factor(xdmgnd$TYPE, labels = c("Derailment", "HeadOn", "Rearend", "Side", "Raking", "BrokenTrain", "Hwy-Rail", "GradeX", "Obstruction", "Explosive", "Fire","Other","SeeNarrative"))
```


Possible predictors of damage	

SPM
Scatter plot matricies for quantitative predictors and single metric.
```{r}
source("SPM_Panel.R")
library(GGally)
ggpairs(xdmgnd[,c("ACCDMG", "TRNSPD", "CARS", "TIMEHR", "TEMP")])
```

PCA
Principal components with the correlation matrix for extreme data with 1 metric and quantitative predictors.
```{r}
source("PCAplots.R")
pred.pca <- princomp(xdmgnd[,c("ACCDMG", "TRNSPD", "CARS", "TIMEHR", "TEMP")], cor = T )
```

Which predictors are most correlated with accident damage?


Categorical plots
heatmap
```{r}
source("http://www.phaget4.org/R/myImagePlot.R")
myImagePlot(table(xdmgnd$Cause, xdmgnd$Type), title = "No. of Accidents by Cause and Type of Accident")
```

Which accident causes and types have the highest numbers of extreme accidents?

Type & TRNSPD
```{r}
library(lattice)
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ Type, scales = "free")
```

Cause & TRNSPD
```{r}
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ Cause, scales = "free")
```

What is notable about the relationship between train speed and accident
damages for different accident causes and types?


Cause X TYPE and TRNSPD
```{r}
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ Cause * Type, scales = "free")
```

Create the Derail variable & 
then look at interactions with Cause
```{r}
xdmgnd$Derail <- (xdmgnd$Type == "Derailment")
```

Plot xy with interactions of Derail and Cause
```{r}
qplot(log(ACCDMG), TRNSPD, data = xdmgnd) +  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + facet_wrap(~ Cause * Derail, scales = "free")
```

Create a Freight variable
then look at interactions with Cause 
```{r}
xdmgnd$Freight <- (xdmgnd$TYPEQ == "1")
```

Interaction plots


Get interaction plot for Derail and Cause
```{r}
ggplot() +
  aes(x = xdmgnd$Derail, y = log(xdmgnd$ACCDMG), group = xdmgnd$Cause, color = xdmgnd$Cause) +
  stat_summary(fun= mean, geom = "point") +
  stat_summary(fun = mean, geom = "line")

```


Interaction plots with quantitative variables
```{r}
Speed <- cut(xdmgnd$TRNSPD, c(min(xdmgnd$TRNSPD),15,max(xdmgnd$TRNSPD)), include.lowest = T, labels = c("low speed", "high speed"))

Cars <- cut(xdmgnd$CARS, c(min(xdmgnd$CARS),1,max(xdmgnd$CARS)), include.lowest = T, labels = c("low hzd", "high hzd"))

Tons <- cut(xdmgnd$TONS, c(min(xdmgnd$TONS),median(xdmgnd$TONS),max(xdmgnd$TONS)), include.lowest = T, labels = c("low tons", "high tons"))

```

Plot interaction between Speed and Cars
```{r}
# First Option with seeing points
qplot(x = TRNSPD, y = log(ACCDMG), data = xdmgnd, colour = Cars) +
  geom_point(colour = "gray")+
  geom_smooth(method = "lm") 

# Second Option

ggplot() +
  aes(x = Speed, y = log(xdmgnd$ACCDMG), group = Cars, color = Cars) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line")

```

Plot interaction between Freight and Speed
```{r}
ggplot() +
  aes(x = xdmgnd$Freight, y = log(xdmgnd$ACCDMG), group = Speed, color = Speed) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line")
```

Plot interaction between Derail and Speed
```{r}
ggplot() +
  aes(x = xdmgnd$Derail, y = log(xdmgnd$ACCDMG), group = Speed, color = Speed) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line")
```

Plot interaction between Tons and Speed
```{r}
ggplot() +
  aes(x = Speed, y = log(xdmgnd$ACCDMG), group = Tons, color = Tons) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.y = mean, geom = "line")
```




How do these results inform your hypotheses?
Use the multivariate visualiztions as evidence to form at least 1 hypothesis.


Linear Models

Build linear regression models with different combinations of quantitative predictors to provide evidence for your hypothesis

Single predictor
```{r}
xdmgnd.lm1<-lm(ACCDMG~TEMP,data=xdmgnd)
summary(xdmgnd.lm1)
names(xdmgnd.lm1)
coef(xdmgnd.lm1)
sum(xdmgnd.lm1$res^2)
```

Two predictors
```{r}
xdmgnd.lm2<-lm(ACCDMG~TEMP+TRNSPD,data=xdmgnd)
summary(xdmgnd.lm2)
names(xdmgnd.lm2)
coef(xdmgnd.lm2)
```

Three predictors
```{r}
xdmgnd.lm3<-lm(ACCDMG~TEMP+TRNSPD+CARS,data=xdmgnd)
summary(xdmgnd.lm3)
coef(xdmgnd.lm3)
```


Interperet your model coefficients.  Do they make sense?



Interperet your developed models using the model utility test and t-test.



Write out the null and alternative hypothesis for each of the tests.  



Do you reject or fail to reject H0 for each test?


Now repeat for TOTKLD + TOTINJ
