---
pdf_document: default
title: "Multiple Linear Regression 4"
abstract: |
  This is the code for Transformations, 
  Qualitative Variables & ANCOVA.
output: pdf_document
---

```{r "setup", include=FALSE}
#This block of code will not appear in your knitted document
#Setup paths for data and Rcode
require("knitr")
traindir <- "~/Google Drive/UVA/Courses/LSM/Fall2020/TrainData/"
sourcedir <-"~/Google Drive/UVA/Courses/LSM/Fall2020/Source/"
opts_knit$set(root.dir = sourcedir)
opts_chunk$set(warning=FALSE)
```

Load data and combine data from 2001-2019 into totacts
```{r}
# Source AccidentInput
source("AccidentInput.R")

# you should have two data structures in working memory
# First - a list of data frames for each year of accident data
acts <- file.inputl(traindir)

# Next a data frame with all accidents from all years from 2001 - 2018
# with columns that are consistent for all of these years

# Get a common set the variables

comvar <- intersect(colnames(acts[[1]]), colnames(acts[[8]]))

# the combined data frame
totacts <- combine.data(acts, comvar)
```

Build a data frame with only extreme accidents for ACCDMG
```{r}
dmgbox <-boxplot(totacts$ACCDMG)
xdmg <- totacts[totacts$ACCDMG > dmgbox$stats[5],]
```

Data cleaning 
```{r}
# Remove 9/11
xdmg <- xdmg[-190,]

##Remove duplicates
xdmgnd <- xdmg[!(duplicated(xdmg[, c("INCDTNO", "YEAR", "MONTH", 
                                     "DAY", "TIMEHR", "TIMEMIN")])),]

# Reset rownames (observation #s) for sequential numbering- 
# otherwise they will remain the #s from totacts
rownames(xdmgnd) <- NULL
```


Build linear regression models in R: lm 

Build 2 multiple linear regression models with the data fram xdmgnd: 
xdmgnd.lm1  ACCDMG ~ TEMP + TRNSPD + TONS + CARS + HEADEND1
xdmgnd.lm2  ACCDMG ~ TEMP + TONS + CARS 

```{r}
xdmgnd.lm1<-lm(ACCDMG ~ TEMP + TRNSPD  + CARS + HEADEND1, data=xdmgnd)
xdmgnd.lm2<-lm(ACCDMG ~ TEMP + CARS , data=xdmgnd)
```

Display regression results for each model
```{r}
summary(xdmgnd.lm1)
summary(xdmgnd.lm2)
```

Create models with interactions between all of the quantitative predictors ----
```{r}
xdmgnd.lm3<-lm(ACCDMG~(TEMP+TRNSPD+CARS+HEADEND1)^2,data=xdmgnd)
summary(xdmgnd.lm3) 
```

Is this the complete second order model?

I() allows your model to contain normal mathematical sysmbols 
Create complete second order model           
```{r}
xdmgnd.lm4<-lm(ACCDMG~(TEMP+TRNSPD+CARS+HEADEND1)^2+I(TEMP^2)+I(TRNSPD^2)+I(CARS^2)+I(HEADEND1^2),data=xdmgnd)
summary(xdmgnd.lm4)
```

How many parameters and coefficients are in each of the models?


Create a stepwise regression model on xdmgnd.lm4
```{r}
xdmgnd.lm4.step <- step(xdmgnd.lm4)
summary(xdmgnd.lm4.step)
```

partial F test
```{r}
anova(xdmgnd.lm3,xdmgnd.lm4)
```

Which model is better based on the partial F test, the larger or smaller?

Interaction Plot Example
```{r}
trnspdbox<-boxplot(xdmgnd$TRNSPD)
TRNSPD.factor<-xdmgnd$TRNSPD
TRNSPD.factor[which(xdmgnd$TRNSPD<50)]<-'low train speed'
TRNSPD.factor[which(xdmgnd$TRNSPD>=50)]<-'high train speed'
TRNSPD.factor <- factor(TRNSPD.factor)
```

Create a derailment variable
```{r}
Derail <- rep(0, nrow(xdmgnd))
Derail[which(xdmgnd$TYPE == 1)] <- 1 
Derail <- as.factor(Derail)
contrasts(Derail)
```

create an interaction plot for TRNSPD and Derailments
```{r}
interaction.plot(TRNSPD.factor,Derail, xdmgnd$ACCDMG)
```



Qualitative Variables
Create a qualitative variable for Cause 
```{r}
Cause <- rep(NA, nrow(xdmgnd))
##Miscellaneous
Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "M")] <- "M"  
##Track
Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "T")] <- "T"  
##Signal or communication
Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "S")] <- "S"   
##Human
Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "H")] <- "H"  
##Electrical or mechanical
Cause[which(substr(xdmgnd$CAUSE, 1, 1) == "E")] <- "E"  
```

# This new variable, Cause, has to be a factor
```{r}
Cause <- as.factor(Cause)
```

# Lets look at the default treatment coding
```{r}
contrasts(Cause)
```

# What is the base case for Cause?



# Write a model to predict ACCDMG in terms of Cause 
```{r}
xdmgnd.lm5<-lm(ACCDMG~Cause,data=xdmgnd)
summary(xdmgnd.lm5) 
```

# How do we interperet the model xdmgnd. lm5?



#Chage based case to H ----
```{r}
contrasts(Cause)<-matrix(c(1,0,0,0,0, 0,0,1,0,0, 0,0,0,1,0,  0,0,0,0,1),nrow=5)
colnames(contrasts(Cause)) <-matrix(c("E","M","S","T"),ncol=4)
contrasts(Cause)
```

```{r}
xdmgnd.lm6<-lm(ACCDMG~Cause,data=xdmgnd)
summary(xdmgnd.lm6)
```

#How do we interperet the model xdmgnd.lm6?


#More qualitative variables:
```{r}
xdmgnd.lm7<-lm(ACCDMG~as.factor(WEATHER),data=xdmgnd)
summary(xdmgnd.lm7) #What is the base case?
```

***************************************** Exercise 1 *****************************************
Recode the qualitative variable Weather using the categories in accident descriptions pdf document



What is the base case?


Build a model with the qualitative variable weather



What is your conclusion?



Build a new model to test the hypothesis: "Accidents of type derailment do not increase the severity of rail road accidents."
 



What is the base case?




What is your conclusion?



ANCOVA           
 

```{r}
attach(xdmgnd) #attach the dataset to the R search path. 
```

# detach(xdmgnd) #remove it from the search path

```{r}
summary(ACCDMG) #now you can use variable names directly
```

Create 2 ANCOVA models:

Build a main effects model Cause + TEMP + TRNSPD +  CARS + HEADEND1
```{r}
xdmgnd.lm1 <-lm(ACCDMG~Cause+TEMP + TRNSPD +  CARS + HEADEND1,data=xdmgnd)
summary(xdmgnd.lm1)
```

Build a main effects + interaction model Cause + TEMP + TRNSPD + CARS + HEADEND1 
```{r}
xdmgnd.lm2<-lm(ACCDMG~(Cause+TEMP + TRNSPD + CARS + HEADEND1)^2,data=xdmgnd)
summary(xdmgnd.lm2)
```

Perform a Partial F Test: xdmgnd.lm1 vs. xdmgnd.lm2
```{r}
anova(xdmgnd.lm1,xdmgnd.lm2)
```

Which model do you choose?


Use stepwise regression on your main effects + interaction model xdmgnd.lm1
```{r}
xdmgnd.lm1.step <- step(xdmgnd.lm2)
```

What is your new model?
```{r}
summary(xdmgnd.lm1.step)
```

**************************************** Exercise 2 *****************************************
Choose 1 of your project 1 hypotheses (or assignment 2) based on an actionable, 
controllable variable.  Build a model with both qualitative
and quantitative variables to test your hypothesis.  Make sure 
to address the following steps:
1.  Variable selection for linear models.
2.  Treatment of categorical variables for your linear model.
3.  Measure the performance of the models.
4.  Adjust your models based on analytical and graphical diagnostics.
5.  Reassess models based on adjustments.
6.  Compare your best models.
7.  Use your best model to reject or accept your hypothesis and provide a 
    recommendation supported by statistical evidence.


